---
title: "Mousetracking Analysis"
author: "Laura Bock Paulsen, Luke Ring"
date: "9/11/2022, updated 26/9/2023"
output: html_document
---



For the analysis of your very own mouse-tracking data, we will be using a package called MouseTrap. In order to get acquainted with the package, we will play around with data from a single participant conduction the circle-square experiment. 

Alternatively, if you feel (ever so slightly) more adventerous, you can also use the data from the touch-tracking experiment. In this case, you will need to use the function "scale_raw_mousetrap_data" to scale the x and y coordinates within subjects. This is because the touch-tracking data needs to be standardized in some way despite differences in people's screen sizes. If you go for this option, you will have an advantage when it comes to analyzing your real data, because it will not be limited to a single subject (but you can ignore the screen size part, which is why it is included as a helper function and not part of the exercise itself). *Hint:* you can use `read_bulk` (from the package `readbulk`) to get all the CSV files from a single directory.

The documentation for the package can be found here: [http://pascalkieslich.github.io/mousetrap/reference/mousetrap.html](http://pascalkieslich.github.io/mousetrap/reference/mousetrap.html).

Use this page to solve the following steps by finding and applying appropriate mousetrap functions. Also, writing ?function_name() in the console is great way of reading about the function in question.

Call me for help whenever you get stuck (or take a quick look in the solutions file) :D

### Install packages and load in the data

*Note:* If you haven't encountered pacman before, it is an easy way to download and load packages for R. You will (only once) need to install packman with `install.packages("pacman")` in R. Afterwards, you can load and install packages with `pacman::p_load(package1, package2, package3)`. If you want to install a package without loading it, you can use `pacman::p_load(package1, package2, package3)`.

```{r setup, include=FALSE}
# loading packages
pacman::p_load(mousetrap, tidyverse, readbulk, png, grid)

```


Touchtracker data:

```{r read_data_tt}
#read in data
d_tt <- read_bulk('../log_files/', 
          extension = 'csv')

### forgot to track actual button click so the following code extracts the last x-position of the mouse, and infers the what button was clicked
#empty list
last_x_pos = c()
number_of_coords = c()

#for each list chars containing x-positios get the last value, turn into nuimeric and append to empty list
for (vector in d_tt$xpos_mouse) {
  number_vector <- (str_extract_all(vector, "-?\\d+\\.?\\d*"))
  numeric_vector <- as.numeric(unlist(number_vector))
  last_x_pos <- append(last_x_pos, tail(numeric_vector, 1))
  
  #also make a list that counts the number of coordinates counted per round
  number_of_coords <- append(number_of_coords, length(numeric_vector))
  
}

#save the list to the dataframe
d_tt$last_x_pos <- last_x_pos

#save the number of coordinates list
d_tt$number_of_coords <- number_of_coords

#add button clicked
d_tt <- d_tt %>% 
  mutate(button_clicked = if_else(last_x_pos < 0, "left_button", "right_button"))

#add whether participant made deontological or utilitarian judgment
d_tt <- d_tt %>% 
  mutate(judgement = if_else(
    button_clicked == deontological_button, "deontological", "utilitarian"
  ))

# #filter for only personal dilemmas
impersonal_dilemmas <- c("\\Standard Trolley\\", "\\Standard Fumes\\")
d_tt <- d_tt %>%
  filter(dilemma_name != impersonal_dilemmas) %>% 
  filter(number_of_coords > 10)


m_tt <- mt_import_mousetrap(
  d_tt,
     verbose = TRUE
)

# now we can remap the trajectories
m_tt <- mt_remap_symmetric(
  m_tt,
  use = 'trajectories',
  remap_xpos = "right",
  remap_ypos = "up"
)

m_tt_tn <- mt_time_normalize(
  m_tt,
  nsteps = 101,
  verbose = TRUE
)

mt_plot_aggregate(
  m_tt_tn, 
  color = "judgement",
  use = 'tn_trajectories'
) +
  ggplot2::scale_color_brewer(type="qual", palette = "Set2") +
  
  theme_classic() +
  
  labs(title = "Average Mouse Trajectory According to Moral Judgement", caption = "Note that coordinates are remapped for different placement of answer buttons, time normalized and aggregated.") +
  
  #xaxis
  xlab("x-position of cursor") +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
  
  #yaxis
  ylab("y-position of cursor") +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
  
  #legend
  labs(color = "Moral judgement") +
  theme(legend.position = c(.95, .05), legend.justification = c("right", "bottom"), legend.box.just = "right", legend.margin = margin(6, 6, 6, 6))

ggsave("average_mouse_trajectory.png")

```




### Mirror-symmetric mapping of movements

Touchtracker data:

```{r plot of answer screen with trajectories overlayed }


mt_plot(
  m_tt, 
  color = "judgement",
  use = 'trajectories'
) +
  ggplot2::scale_color_brewer(type="qual", palette = "Set2") +
  
  theme_classic() +
  
  labs(title = "Raw Mouse Trajectories for Every Trial", caption = "Note that coordinates are remapped to account for the answer buttons randomly switching positions") +
  
  #xaxis
  xlab("x-position of cursor") +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
  
  #yaxis
  ylab("y-position of cursor") +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
  
  #legend
  labs(color = "Moral judgement") +
  theme(legend.position = c(.95, .05), legend.justification = c("right", "bottom"), legend.box.just = "right", legend.margin = margin(6, 6, 6, 6))

ggsave("raw_mouse_trajectories_for_every_trial.png")
  
  

```


```{r}
# an example of aligning the data
m_align <- mt_align(
  m_tt_tn,
  use = 'trajectories',
  dimensions = c("xpos", "ypos"),
  coordinates = c(0,0,-350,250),
  align_start = T,
  align_end = T
)

mt_plot(
  m_align, 
  use = 'trajectories',
  color = "judgement",
  wrap_var = "judgement")

```

```{r}
m_tt_tn <- mt_measures(m_tt_tn, use = 'tn_trajectories')

mt_measures_ag <- mt_aggregate(
  m_tt_tn,
  use_variables = c('MAD', 'xpos_flips','AUC', 'RT'), # if you want all of the measures, exclude this line
  use2_variables = 'judgement'
  )

d_tt$auc <- m_tt_tn$measures$AUC

#density of AUCs
ggplot(d_tt, aes(x=auc, fill = judgement)) +
  labs(fill = "Moral Judgement") +
  geom_density(alpha = 0.75) +
  ggplot2::scale_fill_brewer(type="qual", palette = "Set2") +
  theme_classic() +
  
  labs(title = "Distributions of AUC for Different Moral Judgements") +
  
  #xaxis
  ylab("Density") +
  
  #yaxis
  xlab("Area Under the Curve") +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
  
  #legend
  theme(legend.position = c(.95, .15), legend.justification = c("right", "bottom"), legend.box.just = "right", legend.margin = margin(12, 12, 12, 12)) 


ggsave("auc_distributions.png")

```


